---
title: "Uncertainty Quantification: Bootstrap CIs and DeLong Tests"
subtitle: "Leakage-aware SMO/SVM audit — GSE25055 discovery / GSE25065 external"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Overview

This notebook attaches **uncertainty** to the headline audit metrics. **No new
modeling is performed here**: every value is obtained by resampling the
already-committed out-of-fold predictions. It addresses the reviewer concern
that the pipeline reports point estimates with no confidence intervals and no
significance test for the leakage gap.

The computations are produced by `scripts/07_bootstrap_ci.R` (canonical, using
`pROC` + `PRROC`). A Python cross-check, `scripts/checks/07_bootstrap_ci_crosscheck.py`,
generated the committed tables and agrees with the R script on point estimates
and DeLong results. Full method and reproducibility notes are in
`results/uncertainty/bootstrap_ci_notes.md`.

```{r read}
ci    <- read.csv("../tables/uncertainty/bootstrap_ci.csv", stringsAsFactors = FALSE)
delta <- read.csv("../tables/uncertainty/delta_auroc_prauc_ci.csv", stringsAsFactors = FALSE)
delong<- read.csv("../results/uncertainty/delong_tests.csv", stringsAsFactors = FALSE)
```

## Per-cohort metric confidence intervals

Stratified percentile bootstrap (B = 2000, seed 20260620). The leaky baseline
AUROC/PR-AUC point estimate is the mean across the five repeated-CV runs;
operating-point metrics pool the repeats.

```{r ci-table}
knitr::kable(
  ci,
  col.names = c("Cohort", "Pipeline", "Metric", "Estimate", "95% CI lo",
                "95% CI hi", "B", "Method")
)
```

## Leakage gap (leaky − guarded): paired test

Paired stratified bootstrap on the same resampled samples, with a two-sided
bootstrap p-value.

```{r delta-table}
knitr::kable(
  delta,
  col.names = c("Comparison", "Metric", "Estimate", "95% CI lo", "95% CI hi",
                "Bootstrap p", "B", "Method")
)
```

## DeLong test (per repeat): AUROC, leaky vs guarded

DeLong is run per repeat because the guarded pipeline has a single out-of-fold
score per sample while the leaky baseline has one per repeat.

```{r delong-table}
knitr::kable(
  delong,
  col.names = c("Comparison", "AUROC leaky", "AUROC guarded", "AUROC diff",
                "DeLong z", "DeLong p", "n")
)
```

## Headline interpretation

- **AUROC and operating-point metrics reproduce the manuscript exactly**;
  PR-AUC matches `PRROC::pr.curve(...)$auc.integral` to ~0.001–0.002.
- **The naive-minus-guarded whole-workflow contrast is positive for AUROC and
  uncertain for PR-AUC.** Values are read directly from the regenerated table:
  ΔAUROC = `r delta$point[delta$metric=="delta_auroc"]`
  (`r sprintf("[%.4f, %.4f]", delta$ci_lo[delta$metric=="delta_auroc"], delta$ci_hi[delta$metric=="delta_auroc"])`;
  bootstrap p = `r delta$boot_p[delta$metric=="delta_auroc"]`) and ΔPR-AUC =
  `r delta$point[delta$metric=="delta_pr_auc"]`
  (`r sprintf("[%.4f, %.4f]", delta$ci_lo[delta$metric=="delta_pr_auc"], delta$ci_hi[delta$metric=="delta_pr_auc"])`).
- All confidence intervals are **wide**, driven by the small pCR minority
  (57 of 306 internally; 42 of 182 externally).

**Interpretation.** The two arms also differ in tuning and resampling, so the
contrast is descriptive and does not isolate the causal effect of supervised
feature-selection placement. This is a **methodology / audit** result, not a
clinical biomarker claim.
